---
title: "HW4"
output: html_document
date: "2024-02-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
sclass <- read.csv("sclass.csv")
nbcsurvey <- read.csv("nbc_pilotsurvey.csv")
ebay <- read.csv("ebay.csv")
gasprice <- read.csv("gasprices.csv")
```

**Name: Chester Ashby**

**UTEID: coa398**

**Link: https://github.com/mulletmommy/StatisticalThinking.git**

## Question 1

### Part A
**Claim: Gas stations charge more if they lack direct competition in sight.**
```{r, echo=FALSE}
mean(Price~Competitors, data = gasprice)
#creates bootstrap of gas price vs gas stations with and without direct competition
boot_competition <- do(10000)*diffmean(Price~Competitors, data = mosaic::resample(gasprice))
#gives confidence interval based off boostrap above
confint(boot_competition, level=0.95)
```

**Conclusion: The difference in price for gas stations with and without other gas station brands in sight is somewhere between -.06 and .01. Due to the fact that a mean difference of 0 is within the range of the confidence interval, this indicates that this mean difference is not statistically significant. Therefore the claim is unsupported by the data.**


### Part B
**Claim: The richer the area, the higher the gas prices.**
```{r, echo=FALSE}
lm(Price ~ Income, data=gasprice)
#creates bootstrap of gas price vs median income by ZIP codes
boot_moneyGas = do(10000)*lm(Price ~ Income, data=mosaic::resample(gasprice))
#gives confidence interval based off boostrap above
confint(boot_moneyGas, level = 0.95)
```

**Conclusion: The slope of Gas Prices versus Median Household Income by ZIP Codes is between 6.6e-07 and 1.8e-06, with 95% confidence. This really has no bearing on this data as the difference in gas price is so low that it doesn't support the claim.**


### Part C
**Claim: Gas stations at stoplights charge more.**

```{r, echo=FALSE}
#creates bootstrap of gas price vs whether gas station exist at a stoplight
boot_stoplight <- do(10000)*diffmean(Price~Stoplight, data = mosaic::resample(gasprice))
#gives confidence interval based off boostrap above
confint(boot_stoplight, level=0.95)
```

**Conclusion: The mean difference in price for gas stations with and without other gas station brands in sight is somewhere between -.04 and .03 dollars with 95% confidence. Due to the fact that a mean difference of 0 is within the range of the confidence interval, this indicates that this mean difference is not statistically significant. Therefore the claim is unsupported by the data.**


### Part D
**Claim: Gas stations with direct highway access charge more.**

```{r, echo=FALSE}
#creates bootstrap of gas price vs whether gas station is accesisble from highway
boot_highway <- do(10000)*diffmean(Price~Highway, data = mosaic::resample(gasprice))
#gives confidence interval based off boostrap above
confint(boot_highway, level=0.95)
```

**Conclusion: The mean difference in price for gas stations with and without highway access is somewhere between .01 and .08 dollars, with 95% confidence. This mean difference is statistically significant at the 5% (0.05) level. This means that the claim is supported by the data.**


### Part E
**Claim: Shell charges more than all other non-Shell brands.**

```{r, echo=FALSE}
shell_gasprice <- gasprice %>% 
  mutate(isShell = ifelse(Brand == "Shell", TRUE, FALSE))
#creates bootstrap of gas price vs whether gas station's brand is shell or not
boot_shell <- do(10000)*diffmean(Price~isShell, data = mosaic::resample(shell_gasprice))
#gives confidence interval based off bootstrap above
confint(boot_shell, level=0.95)
```

**Conclusion: The mean difference in price for gas stations vs brands that were or weren't Shell is somewhere between -.01 and .06 dollars with 95% confidence. Due to the fact that a mean difference of 0 is within the range of the confidence interval, this indicates that this mean difference is not statistically significant. Therefore the claim is unsupported by the data.**


## Question 2

```{r, echo=FALSE}
#creates new dataset filtering year 2011 and trim of 63 AMG
edit1_sclass <- sclass %>% 
  filter(year == 2011, trim == "63 AMG")
#creates bootstrap of the mean mileage of the dataset edit1_sclass
boot_edit1_sclass <- do(10000)*mean(~mileage, data = mosaic::resample(edit1_sclass))
#gives confidence interval based off bootstrap above
confint(boot_edit1_sclass, level=0.95)
```


```{r, echo=FALSE}
#creates new dataset of only Mercedes of year 2014 and trim of 550, create new variable isBlack
edit2_sclass <- sclass %>% 
  filter(year == 2014, trim == "550") %>% 
  mutate(isBlack = ifelse(color == "Black", TRUE, FALSE))

prop(~isBlack, data=edit2_sclass)
#creates bootstrap of proportion of cars that are Black vs isnt Black
boot_edit2_sclass <- do(10000)*prop(~isBlack, data = mosaic::resample(edit2_sclass))
#gives confidence interval based off bootstrap above
confint(boot_edit2_sclass, level=0.95)
```


## Question 3


### Part A
**Question: Who makes people happier: Ed or Earl?**
**Approach: I used the diffmean() function to find difference in mean viewer response to the Q1_Happy question for these two shows.**

```{r, echo=FALSE}
#creates new dataset with only 2 shows
happynbcsurvey <- nbcsurvey %>% 
  filter(Show == "Living with Ed" | Show == "My Name is Earl")
#creates bootstrap of happy ratings vs show
boot_happynbcsurvey = do(10000)*diffmean(Q1_Happy ~ Show, data=mosaic::resample(happynbcsurvey))
#gives confidence interval based off bootstrap above
confint(boot_happynbcsurvey, level = 0.95)
```

**Conclusion: The mean difference in viewer Happiness vs the 2 different shows is somewhere between -.39 and .10 points with 95% confidence. Due to the fact that a mean difference of 0 is within the range of the confidence interval, this indicates that this mean difference is not statistically significant. Therefore there is no evidence that one show consistently produces a higher mean Q1_Happy response among viewers?**


### Part B
**Question: Which reality/contest show made people feel more annoyed? The Biggest Loser or The Apprentice: Los Angeles.**
**Approach: I used the diffmean() function to find difference in mean viewer response to the Annoyance question for these two shows.**
```{r, echo=FALSE}
#creates new dataset of only the 2 shows below
annoynbc <- nbcsurvey %>% 
  filter(Show == "The Biggest Loser" | Show == "The Apprentice: Los Angeles")
#creates bootstrap of annoyance rating of both shows
boot_annoy = do(10000)*diffmean(Q1_Annoyed ~ Show, data=mosaic::resample(annoynbc))
#gives confidence interval based off bootstrap above
confint(boot_annoy, level = 0.95)
```

**Conclusion: The mean difference in viewer annoyance vs the 2 different shows is somewhere between -.52 and -.01 points with 95% confidence. This mean difference is statistically significant at the 5% (0.05) level. There is evidence that one show consistently produces a higher mean Q1_Annoyed response among viewers.**


### Part C
**Question: What proportion of American TV watchers would we expect to give a response of 4 or greater to the “Q2_Confusing” question?**
**Approach: I used the prop() function to find the proportion of American TV watchers that we expect to give a response of 4 or greater to the “Q2_Confusing” question.**
```{r, echo=FALSE}
#creates bew dataset of only 1 show and creates new is4Plus variable
confuse_nbc <- nbcsurvey %>% 
  filter(Show == "Dancing with the Stars") %>% 
  mutate(is4Plus = ifelse(Q2_Confusing >= 4, TRUE, FALSE))
#creates bootstrap of proportions for is4Plus variable
boot_confusion <- do(10000)*prop(~is4Plus, data = mosaic::resample(confuse_nbc))
#gives confidence interval based off bootstrap above
confint(boot_confusion, level=0.95)
```

**Conclusion: The proportion of American TV watchers that we expect to give a response of 4 or greater to the “Q2_Confusing” question is somewhere between 3.87 and 11.60 percent with 95% confidence.**


## Question 4

**Question: Is the revenue ratio is the same in the treatment and control groups, or whether instead the data favors the idea that paid search advertising on Google creates extra revenue for EBay?**
**Approach: I used the diffmean() function to find difference in mean revenue ratio between paid searches or organic searches.**

```{r, echo= FALSE}
#creates new dataset of rev_ratio vs adwords_pause
boot_ebay <- do(10000)*diffmean(rev_ratio ~ adwords_pause, data=mosaic::resample(ebay))
#gives confidence interval based off bootstrap above
confint(boot_ebay)
```


**Conclusion: The mean difference in mean revenue ratio between paid searches or organic searches is somewhere between -.09 and -.01 percent with 95% confidence. This mean difference is statistically significant at the 5% (0.05) level. This shows that the paid search advertising were a waste of money, as the control groups revenue ratio was higher. **

