---
title: "HW5"
output: html_document
date: "2024-02-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
```

**Name: Chester Ashby**
**UTEID: coa398**
**Link: https://github.com/mulletmommy/StatisticalThinking.git**


## Question 1

**Null Hypothesis: Over the long run, securities trades from the Iron Bank are flagged at the same 2.4% baseline rate as that of other traders.**

**Test Statistic: number of flagged trades that are over 70.**

```{r, echo=FALSE}
sim_flagtrades = do(100000)*nflip(n=2021, prob=.024)


ggplot(sim_flagtrades) + 
  geom_histogram(aes(x=nflip), binwidth=1)

p_value <- sum(sim_flagtrades >= 70)/100000
round(p_value, 3)
```

**Since the p-value score is 0.002, which is less than 0.05, it indicates that the p-value is statistically significant. In this case of flagged trades from Iron Bank, there is a possibility that there could be Insider Trading or something else that could explain the p-value.**


## Question 2


**Null Hypothesis: The number of health code violations by Gourmet Bite are consistent with the average 3% baseline rate that restaurants in the city are cited for.**

**Testing Statstic: number of health code violations that are over 8.**

```{r, echo=FALSE}
sim_violation = do(100000)*nflip(n=50, prob = 0.03)


ggplot(sim_violation) +
  geom_histogram(aes(x=nflip), binwidth=1)

p_value2 <- sum(sim_violation >= 8)/100000
round(p_value2, 3)
```


**The p-value score is 0.000, which is less than 0.05, and therefore indicates that the p-value is statistically significant. It is highly likely that Gourmet Bite are cited for health code violations at higher rate than the average 3% baseline rate of the city. Therefore the Health Department should take action against Gourmet Bites.** 


## Question 3

```{r, echo=FALSE}
brown_sentences <- readLines("brown_sentences.txt")

# Import letter frequencies
letter_frequencies = read.csv("letter_frequencies.csv")

# This will calculate the chi-squared goodness of fit statistic
# for an input sentence, based on a frequency table of letters
calculate_chi_squared = function(sentence, freq_table) {
  
  # Ensure letter frequencies are normalized and sum to 1
  freq_table$Probability = freq_table$Probability / sum(freq_table$Probability)
  
  # Remove non-letters and convert to uppercase
  clean_sentence = gsub("[^A-Za-z]", "", sentence)
  clean_sentence = toupper(clean_sentence)
  
  # Count the occurrences of each letter in the sentence
  observed_counts = table(factor(strsplit(clean_sentence, "")[[1]], levels = freq_table$Letter))
  
  # Calculate expected counts
  total_letters = sum(observed_counts)
  expected_counts = total_letters * freq_table$Probability
  
  # Chi-squared statistic
  chi_squared_stat = sum((observed_counts - expected_counts)^2 / expected_counts)
  
  return(chi_squared_stat)
}

##creates empty vector for all the chi_squared values of brown_sentences
chi_sq = rep(0, length(brown_sentences))

#loops through brown_sentences and adds the its chi_squred value to chi_sq
for (i in 1:length(brown_sentences)){
  chi_sq[i] = calculate_chi_squared(brown_sentences[i], letter_frequencies)
}

#plotting chi_sq
plot(chi_sq, log='y')

```


```{r, echo = FALSE}
sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)
#creating empty vector for "sentences"'s chi_squared values
the_one <- rep(0, 10)

#looping through "sentences" and adding its following chi_squared values to "the_one"
for (i in 1:10){
  chi_value = calculate_chi_squared(sentences[i], letter_frequencies)
  the_one[i] = sum(chi_sq >= chi_value)/sum(chi_sq)
}

#making a tibble of "the_one"
the_one_tibble <- tibble(the_one) %>% 
  mutate(sentence = 1:10) %>% 
  mutate(p_value = round(the_one, 3)) %>% 
  select(sentence, p_value)

the_one_tibble
```


**The sentence water marked by an LLM is sentence 6. It has a p-value of 0.000, which is way smaller than 0.05, which indicates that the p-value is statistically significant.Sentence 6's is extremely close to actual 0, which means that it is unlikely to be described by chance, but rather by a cause, which in this case is the watermark.**

